\textsc{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}

### 2018-02-23 redaction/dhp_july2017.tex
Les changements de Jérôme

Trade costs are central to international economic analysis. In particular, they are considered a major obstacle to international economic integration and international trade flows. After the Second World War though, the reduction of policy trade costs have been the workhorse of the GATT (WTO after 1995) rounds of multilateral negotiations. Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000. Despite this, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\cite{Head_Mayer04} or \cite{Disdier_Head08}, to name a few). Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs would represent a 21\% markup over production costs. Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) indicates that the impact of transport costs on trade flows is not negligible. If much of trade policy barriers have been removed over the second half of the 20th century, this suggests that the transport costs component remains large and deserves attention. International transport costs accordingly stand at the heart of the paper.\smallskip

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others). Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). %According to \cite{Glaeser04}, over the twentieth century, the cost of moving goods inland have declined by over 90\% in real terms.
The container was a major transport-cost reducing innovation (see \citealp{levinson2016box}). However, \cite{hummels2007} points out the role of trade composition effects in accounting for the time trends of international transport costs. Based on US imports over 1974-2004, he finds that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have decreased in both air and maritime shipment. However, he also finds that trade composition effects partially compensate this ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend. In this paper, we investigate the robustness of this result. As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, the marked reduction of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport. In maritime shipment, trade composition effects matter much, but they amplify (rather than offset) the reduction in pure transport costs.\medskip

To prove this point we provide a careful modelling of transport costs. Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see the references below). Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs. This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive parts. To do so, we exploit information contained in the US imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis. This database is the same as \cite{hummels2007}, extended over the more recent years (2005-2013). \smallskip

@@

Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se''. Trade composition effects matter much, and when they matter (in maritime transport), they amplify the reduction of ``pure'' transport costs, strengthening the downward trend for overall transport costs. This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer. Allowing the additive costs to vary  across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. Our result thus stands in line with common knowledge observations of the reduction of transport costs due, e.g., to the diffusion of the container.\smallskip

@@

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Based on Norwegian firm-level export data in 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the ad-valorem trade costs (expressed in Norwegian crowns). Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive to the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value of 2.7 USD paid in additive costs for maritime transport and 1.4 USD for air shipping. This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode). \footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

Our analysis of transportation costs consists in exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007} (among others). The database we use to construct our measure of transport costs comes from US annual Imports of Merchandise provided by the Census bureau, spanning from 1974 to 2013. Details on the database are provided in Appendix \ref{app:data}. We first use customs values, quantities and freight costs to recover free-alongside (fas) and cost-insurance-fret (cif) prices, by goods, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob price rather than the fas price. The fas price (for ``Free Alongside Ship'') means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in. It is also the seller's responsibility to clear the goods for export. The fob price (for ``Free on Board'') means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer. Once the goods clear the railing of the vessel the buyer assumes the risk. Note that this term is used exclusively for maritime and inland waterway transport. While both terms are closely related, we refer to the fas price as the price considered by the US Foreign Trade Statistics.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs. The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight. Our dependant variable is finally computed as the ratio of the cif price divided by the fas price. Higher than 1, the variable provides a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent. This is a quite standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.

@@

Exploiting this dataset also has (at least) three advantages. First, it delivers a strong statistical reliability arising from a single, homogenous and trustworthy customs source.  This limits the measurement error bias. Based on customs declarations, the US Imports database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system. This will be crucial to compute transport costs, as we detail below. In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, ocean (or ``vessel'') and air. We will make use of this to enlighten potential differences in the dynamics of transport costs across transportation mode. Second, using this dataset allows us to have both the import price and the export price for the same good (for a given origin country). This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to \cite{Irrazabal_2015}. Third, this dataset is available over a long time span, which we exploit by providing an analysis of the trend patterns of international transport costs over the period 1974-2013. Notice that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e. ignoring those flows that did not happen because of presumably prohibitive transport costs. In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs.\smallskip

We estimate international transport costs at the 3-digit classification level, even if data series on the cif and fas prices are available at the 5-digit classification level. As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013. Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation. However, we obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels. Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the considered year, this leaves us with around 200 sectors at the 3-digit level, from approximately 200 countries of origin.

@@

As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international trade costs, that decomposes into both an additive and an ad-valorem component. In this respect, our paper relates to the recent literature about the importance of additive costs in accounting for international trade costs (\cite{Irrazabal_2015}, among others). We contribute to this issue by adopting the following strategy. For each year and transport mode, we estimate two models: \textit{(a)} when additive costs are excluded, in which case transport costs are modeled as iceberg costs only (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})). From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modelling the additive component.

@@

After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode. When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$. In this case, notice that the equation could be estimated relying on a linear form. To preserve comparability of the results, we keep the same non-linear estimation method in both cases though. Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme. We thus recover a ``synthetic estimate'' of each type of transport cost: \textit{a)} $\widehat{\tau}^{ice}$, \textit{b)} $\widehat{\tau}^{adv}$ and $\widehat{t}$, for each year and transportation mode. These results are reported in Section \ref{sec:results_decomposition}.

@@

Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components. The first step of analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})). This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of transport costs. In this respect, these results \textit{per se} also constitute our first original contribution to the literature. In Appendix \ref{app:diagnostic_test}, we provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation.\smallskip

@@

Coming now to interpretation, two main results emerge from Table \ref{tab:summary_results}. First, it provides estimation results about the size of the overall transport costs. From Table \ref{tab:summary_results}, we obtain an overall value of approximatively a 6\% markup in ocean shipping, 5\% markup in air shipping in average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$. Under Model (A) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period). Under Model (B), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$. With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in Vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in Vessel. A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{The difference of results with \citet{anderson_wincoop_jel} can be accounted for by a number of reasons. \citet{anderson_wincoop_jel} obtain this 11\% markup as a rough value by aggregating various sources over a bunch of industrialized countries, with data mostly based on the second half of the 1990s. By contrast, our estimation of a 6\% markup relies on a single source of data based on the US imports flows over 1974-2013. Beyond country coverage and the time sample, another plausible candidate explanation lays in the difference of empirical methodology: \citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}

Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs. On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.12 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case. Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8 \% for ocean shipping and air transport respectively. This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 1.8\% for vessel and air shipping respectively (on average over the period). Put it differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air. This shows that the per-kg cost dimension is quantitatively sizeable. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2. Additionally, we run quality-of-fit diagnostic tests that confirms the importance of the additive cost component. The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degree of freedom, as reported in Appendix \ref{app:diagnostic_test}. Further, these results provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.\smallskip

>>>>

Trade costs are central to international economic analysis. In particular, they are considered a major obstacle to international economic integration and international trade flows. After the Second World War though, the reduction of policy trade costs have been the workhorse of the GATT (WTO after 1995) rounds of multilateral negotiations. Based on panel data, \citet{novy13} finds  that U.S. trade costs with major trading partners declined on average by about 40 percent between 1970 and 2000. Despite this, several papers (mostly based on empirical estimates of the gravity equation) have shown that trade costs still remain a major obstacle to trade (\citealp{Head_Mayer04} or \citealp{Disdier_Head08}, to name a few). Using data over 1989-2000, \citet{anderson_wincoop_jel} thus estimate that average international trade costs for industrialized countries represent a 74\% markup over production costs.

Defined as the costs associated with the exchange of goods across national borders, trade costs are usually split between transaction costs (information costs, contract enforcement costs, costs associated with the use of different currencies...), policy costs (tariff  and non-tariff costs), time costs (time to ship goods) and transport costs \emph{per se}. In this vein, \citet{anderson_wincoop_jel} obtain that around 30\% of international trade costs are attributable to transport costs. Equivalently, international transportation costs would represent a 21\% markup over production costs. Further, the sizeable elasticity of trade with respect to freight costs obtained by \cite{Behar_Venables} (around -3) indicates that the impact of transport costs on trade flows is not negligible. If much of trade policy barriers have been removed over the second half of the 20\textsuperscript{th} century, this suggests that the transport costs component remains large and deserves attention. International transport costs accordingly stand at the heart of the paper.\smallskip

This focus is shared with part of the international trade literature that studies the patterns of transport costs over time, such as \cite{hummels2007} and \cite{Behar_Venables} (among others). Many argue that transport costs have substantially decreased with technological advance in transportation, infrastructure development and new communication technologies (see \citealp{Lafourcade_Thisse}). %According to \cite{Glaeser04}, over the twentieth century, the cost of moving goods inland have declined by over 90\% in real terms.
The container was a major transport-cost reducing innovation (see \citealp{levinson2016box}). However, \cite{hummels2007} points out the role of trade composition effects in accounting for the time trends of international transport costs. Based on US imports over 1974-2004, he finds that transport costs ``per se'' (i.e., \textit{ceteris paribus} for given goods, mode, and trade routes) have decreased in both air and maritime shipment. However, he also finds that trade composition effects partly offset this ``pure'' transport costs decrease, thereby attenuating the overall transport costs downward trend. In this paper, we investigate the robustness of this result. As a major contribution, and in contrast to \cite{hummels2007}, we find that trade composition effects play a minor role in accounting for the trend patterns of transport costs. Rather, the marked reduction of transport costs ``per se'' constitutes the main driving force behind the overall transport costs decrease, in particular for air transport. In maritime shipment, trade composition effects matter more, but they amplify (rather than offset) the reduction in pure transport costs.\medskip

To prove this point we provide a careful modelling of transport costs. Following \citet{samuelson1954}, standard models of international trade have usually modeled trade costs as an \emph{ad-valorem} tax equivalent (ie, as a constant percentage of the producer price per unit traded, part of the ``iceberg cost'' hypothesis).\footnote{Strictly speaking, the ``iceberg cost'' includes both the ad-valorem dimension of trade cost and the fact that these costs are paid in terms of the good that is traded. As this last element is irrelevant in our case, we use the terms ``iceberg'' and ``ad-valorem'' interchangeably, as it is commonly done in the literature.} However, a recent strand of the literature points out to the existence of an additive component, that is, a cost per unit traded (see, among others, \citealp{Irrazabal_2015} or \citealp{martin2012}). Based on US sectoral data, our own estimates also bring out the importance of additive costs in international transport costs. This is obtained by providing an empirical decomposition of the structure of transport costs over time, in which we explicitly distinguish between ad-valorem and additive parts. To do so, we exploit information contained in the US imports flows over 1974-2013, from which we get a measure of international transport costs as the difference between the import and the export prices, on a year-product-partner country basis. This database is the same as \cite{hummels2007}, extended over more recent years (2005-2013). \smallskip

@@

Second, we find that the main source of the downward trend in overall transport costs (that decreased by roughly 50\% in air shipping and 60\% in vessel shipping over 1974-2013) does come from the reduction of transport cost ``per se''. Trade composition effects only have a limited influence, and when they matter (in maritime transport), they amplify the reduction of ``pure'' transport costs, strengthening the downward trend for overall transport costs. This stands in sharp contrast to the conclusion obtained by \cite{hummels2007} on the same database (until 2004). Importantly, we show that the difference of results can be attributed to the new way of modeling of the additive component we offer. Allowing the additive costs to vary  across time, products and country partners (rather than being constant as in \citealp{hummels2007}), turns out to be key in the decomposition of underlying sources of the decrease of overall transport costs observed over the period. Our result thus stands in line with common knowledge observations of the reduction of transport costs due, e.g., to the diffusion of the container.\smallskip

@@

Closely related to our paper is the work by \citet{Irrazabal_2015}, which develop a structural framework for inferring additive trade costs from firm-level trade data. Implementing their methodology on Norwegian firm-level export data for the year 2004, they find that, for the median shipment, additive costs (in Norwegian crowns) amount to 6\% of the ad-valorem trade costs (expressed in Norwegian crowns). Our results share in common with \citet{Irrazabal_2015} the important role of the additive component of international trade costs. Yet, our paper complements their findings in two main respects. First, while our study covers a narrower set of trade costs focused on transport costs, their data and their empirical approach only allows the identification of the ratio of the additive to the ad-valorem cost; by contrast, our estimation strategy enables us to uncover separately both values of the ad-valorem and the additive costs. From this, we can rebuilt the ratio in similar terms to them, thereby gaining in generality in this respect. While \citet{Irrazabal_2015} obtain that, for an import price of 100 USD, 6 USD are paid in additive trade costs, our own results point to a value of 2.7 USD paid in additive transport costs for maritime shipping and 1.4 USD for air transport. This comparison of results suggests that between a third and a half of the additive trade costs are attributable to the transport sector (depending on the transport mode). \footnote{See Section \ref{sec:results_decomposition} for more details.} Second, we exploit exhaustive information about the imports flows of the US over a large time span from 1974 to 2013. In this respect, our results deliver a broader view of the magnitude of additive costs in international trade over time. In particular, we show that the modeling of additive costs is of critical importance in determining the underlying sources of the trends patterns of international transport costs observed since 1974.\footnote{On top of the previously cited papers about additive costs, our paper also relates to \citet{Kropf-Saure-JIE-2016}, which estimate the size and shape of per-shipment costs bases on Swiss data, and to \citet{Alessandria-et-al-AER-2010}, \citet{Hornok-et-al-RES-2015} and \citet{Hornok-et-al-JIE-2015}, which point out the role of per-shipment costs (among which, administrative costs) in generating some ``lumpiness'' in international trade transactions.} \smallskip

@@

Our analysis of transportation costs consists in exploiting the difference between commodity-level export and import prices, as in \cite{hummels2007} (among others). The database we use to construct our measure of transport costs comes from US annual Imports of Merchandise provided by the Census bureau, spanning from 1974 to 2013. Details on the database are provided in Appendix \ref{app:data}. We first use customs values, quantities and freight costs to recover free-alongside (fas) and cost-insurance-fret (cif) prices, by goods, country of origin and transportation mode.\footnote{The related literature commonly refers to the fob price rather than the fas price. The fas price (for ``Free Alongside Ship'') means that the seller must transport the goods all the way to the dock, close enough to be reached by the crane of the ship it will be transported in. It is also the seller's responsibility to clear the goods for export. The fob price (for ``Free on Board'') means that the seller is obligated to bring the goods all the way to the port, clear the goods for export, and see that they are loaded onto the ship nominated by the buyer. Once the goods clear the railing of the vessel the buyer assumes the risk. Note that this term is used exclusively for maritime and inland waterway transport. While both terms are closely related, we refer to the fas price as the price considered by the US Foreign Trade Statistics.} More precisely, the (unit) fas price is computed as the total ``customs value'' in the US trade statistics divided by the shipping weight; in other words, it is the price for one kg of the good net of transportation costs. The cif price is then computed as the sum of the customs value and freight charges, once again divided by the shipping weight. Our dependant variable is finally computed as the ratio of the cif price divided by the fas price. By construction higher than 1, the variable provides a measure of transport costs as a proportion of the good's price, an \emph{ad-valorem} equivalent. This is a quite standard and widespread strategy, as emphasized by \citet{anderson_wincoop_jel}.

@@

Exploiting this dataset also has (at least) three advantages. First, it delivers a strong statistical reliability arising from a single, homogenous and trustworthy customs source.  This limits the measurement error bias. Based on customs declarations, the US Imports database inventories all imports (both values and quantities), by country of origin, to the United states at the HS 10-digit level, with a concordance code to the SITC 5-digit coding system. This will be crucial to compute transport costs, as we detail below. In addition, the database reports information regarding freight, insurance and handling expenditures by transportation mode, ocean (or ``vessel'') and air. We will make use of this to enlighten potential differences in the dynamics of transport costs across transportation mode. Second, using this dataset allows us to have both the import price and the export price for the same good (for a given origin country). This is highly valuable, as it gives us a direct measure of international transport costs (at the country/sector level), from which we can estimate separately the \textit{levels} of both the iceberg costs and of the additive costs, in contrast to \cite{Irrazabal_2015}. Third, this dataset is available over a long time span, which we exploit by providing an analysis of the trend patterns of international transport costs over the period 1974-2013. Note that, as mentioned by \cite{Lafourcade_Thisse}, our transport costs measures is based on actual trade flows, i.e. ignoring those flows that did not happen because of presumably prohibitive transport costs. In light of this, the estimated values that we obtain can be viewed as the lower bounds of the ``true'' transport costs.\smallskip

We estimate international transport costs at the 3-digit classification level, even if data series on the cif and fas prices are available at the 5-digit classification level. As detailed below, the use of a nonlinear estimator triggers computational limitations that limit the level of possible detail, especially when covering a long period of time. Yet, we ensure the robustness of these results by conducting the estimations at the 4-digit level for some selected years.\footnote{The selected years for the 4-digit level estimations are: 1974, 1977, 1981, 1985, 1989, 1993, 1997, 2001, 2005, 2009, 2013. Comparing different levels of aggregation is useful to check differences and the presence of biases precisely due to aggregation. However, we obtain no substantial difference between the estimation results conducted at the 3 and 4-digit levels. Estimation results at the 4-digit classification level are reported in Appendix \ref{app:4digit}.} Depending on the considered year, this leaves us with around 200 sectors at the 3-digit level, from around 200 countries of origin.

@@

As mentioned in the Introduction, one contribution of the paper is to provide a careful estimation of international transport costs, that decomposes into both an additive and an ad-valorem component. In this respect, our paper relates to the recent literature about the importance of additive costs in accounting for international trade costs (\citealp{Irrazabal_2015}, among others). We contribute to this issue by adopting the following strategy. For each year and transport mode, we estimate two models: \textit{(a)} when additive costs are excluded, in which case transport costs are modeled as iceberg costs only (Equation (\ref{eq:model_nlI})), and \textit{(b)} when transport costs are decomposed in the two additive and ad-valorem dimensions (Equation (\ref{eq:model_IetA})). From this, we compare the fitting properties and the explanatory power of both models to evaluate the importance of modelling the additive component.

@@

After estimating Equation (\ref{eq:model_IetA}), we can re-built a measure of each component, $\widehat{\tau}^{adv}_{is} = \widehat{\tau_{i}} \times \widehat{\tau}_{s(k)}$ for the ad-valorem cost and $\widehat{t}_{is} = \widehat{t}_{i} + \widehat{t}_{s(k)}$ for the additive cost, that are country-sector specific, by year and transport mode. When assuming iceberg costs only (Equation (\ref{eq:model_nlI})), we proceed similarly to get $\widehat{\tau}^{ice}_{is} = \widehat{\tau}_{i} \times \widehat{\tau}_{s}$. In this case, notice that the equation could be estimated relying on a linear form. To preserve comparability of the results, we keep the same non-linear estimation method in both cases though. Similarly as \citet{Irrazabal_2015}, we take the average over the sector-country dimension, using the values of each trade flow ($is$-specific) over total yearly trade as a weighting scheme. We thus recover a ``synthetic estimate'' of each type of transport cost: $\widehat{\tau}^{ice}$ for model \emph{(a)}, $\widehat{\tau}^{adv}$ and $\widehat{t}$ for model \textit{b)}, for each year and transportation mode. These results are reported in Section \ref{sec:results_decomposition}.

@@

Our ambition to re-assess the underlying sources of the trend patterns of international transport costs requires a careful modeling of both the additive and the multiplicative components. The first step of the analysis then consists in providing a quantification of the magnitude of transport costs over time (by transport mode), distinguishing whether the additive component $t_{ik}$ is excluded or included in the estimated equation (Equation (\ref{eq:model_IetA}) or (\ref{eq:model_nlI})). This allows us to provide estimates for the size of both the ad-valorem and the additive components of transport costs, and assess whether additive costs represent a sizeable component of the latter. In this respect, these results \textit{per se} also constitute our first original contribution to the literature. In Appendix \ref{app:diagnostic_test}, we provide diagnostic tests about the goodness-of-fit measures, that confirm the empirical relevance of adding the additive component in the estimation.\smallskip

@@

Coming now to interpretation, two main results emerge from Table \ref{tab:summary_results}. First, it provides estimation results about the size of the overall transport costs. From Table \ref{tab:summary_results}, we obtain an overall value of approximatively a 6\% markup in ocean shipping, 5\% markup in air shipping on average over 1974-2013.\footnote{Expressed as a markup over production costs, transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}}$. Under Model (A) (ad-valorem costs only), making use of the figures reported in Table \ref{tab:summary_results} gives a markup of 5.8\% in vessel, and 5.1\% in air shipping (considering the mean value over the period). Under Model (B), total transport costs are equal to $\frac{p-\widetilde{p}}{\widetilde{p}} = (\tau-1) + \frac{t}{\widetilde{p}}$. With $\tau-1= 0.032$ and $\frac{t}{\widetilde{p}}= 0.029$ in Vessel (as reported in Table \ref{tab:summary_results}), we get a value equal to 6.1\% in Vessel. A similar calculus gives 4.3\% in air transport.} If this seems modest in comparison with the 11\% of markup obtained by \citet{anderson_wincoop_jel} over a set of industrialized countries, this stands in line with the results obtained by \citet{hummels2007} for the US economy.\footnote{Beyond country coverage and time period, another plausible candidate explanation lays in the difference of empirical methodology: \citet{anderson_wincoop_jel}'s estimates are based on standard gravity equations, while ours come from a non-linear estimation of a ``structural'' decomposition of the difference between export and import prices.}

Second, and most importantly, Table \ref{tab:summary_results} provides new quantitative evidence about the size of the additive component of transportation costs. On average over the whole period, for sectors defined at the 3-digit level, additive costs amount to 10 cents (0.1 USD) per kilogram exported in maritime transport, and 1.12 USD per kilogram exported in air transport, indicating that freight rates are much higher in the latter case. Expressed in terms of the export price, the additive costs amount to 2.9\% and 1.8 \% for ocean shipping and air transport respectively. This is slightly lower than the estimates for the ad-valorem component, which is estimated to 3.2\% and 2.5\% for vessel and air shipping respectively (on average over the period). Put differently, the additive component amounts to 48.2\% of total transport costs in Vessel, and 42.3\% in Air. This shows that the per-kg cost dimension is quantitatively sizeable. For both transport modes, the omission of the additive term seriously biases the ad-valorem term upward, by roughly a factor 2. Additionally, we run quality-of-fit diagnostic tests that confirms the importance of the additive cost component. The goodness of fit is systematically better when the additive component is included in the regression, even when taking into account the additional degree of freedom, as reported in Appendix \ref{app:diagnostic_test}. Further, these results provide quantitative estimates for the values of both additive and ad-valorem transport costs, which can be very useful for more theoretical approaches needing calibration.\smallskip

Our estimation results about the size of additive transport costs can be compared to \citet{Irrazabal_2015}, which constitute the most recent evidence on this question. Based on Norwegian trade data for 2004, \citet{Irrazabal_2015} estimate that additive costs (in Norwegian crowns) amount to 6\% of ad-valorem trade costs (expressed in Norwegian crowns), i.e. $\frac{t}{\tau\widetilde{p}}=6\%$ expressed in our terminology (with $\tau>1$, and retaining their estimated weighted mean value, which is the one directly comparable to our own results). One important difference with respect to their work is that we can provide separate estimates for each cost component (additive and ad-valorem). From this, we can rebuilt the ratio in similar terms to them, allowing us to gain in generality in this respect. Making use of our estimates for 2004, we thus obtain a ratio $\frac{t}{\tau\widetilde{p}}=2.8\%$ in ocean shipping, 1.4\% in air transport. This may sound surprisingly low, in contrast to the 6\% obtained by \citet{Irrazabal_2015}. This can be accounted for by recalling the difference in the type of dataset - hence, of costs, embraced in each case. While the database of \citet{Irrazabal_2015} allows studying trade costs in general, our database covers a subset made of the monetary international transport costs, as we start from the gap between the import and the export prices (on top of differences due to the country coverage of the two databases). In this respect, our two papers can be used in a complementary fashion to infer the proportion of additive trade costs that are due to international shipment. This obeys the following reasoning. \citet{Irrazabal_2015}'s results imply that, for a import price of 100 USD, 6 USD are paid in additive trade costs. For the same import price of 100 USD, our results yield the value of 2.7 (1.4) USD paid in additive transport costs for maritime (air) transport.\footnote{In both papers, starting from the estimated value of the ratio $\frac{t}{\tau \widetilde{p}}= x$, we can rebuilt the proportion of additive costs in terms of the import price $\frac{t}{\tau \widetilde{p} + t} = \frac{x}{1+x}$. With $x = 0.028$ and $0.014$ in vessel and air transport in 2004, we get a ratio $\frac{t}{\tau \widetilde{p} + t}$ equal to 0.027 and 0.014 in each transport mode respectively.} This comparison of results suggests that between 25\% (in Air) and 48\% (in Vessel) of additive trade costs are attributable to international transport. On top of our results summarized in Table \ref{tab:summary_results}, this points out the evidence that the additive component represents a quantitative sizeable dimension of international transport costs. In the next section, we keep on investigating the importance of the (varying pattern of) additive costs, as regards with the determinants of transport costs trend patterns.

\section{Transport costs time trends: The role of the additive component}\label{sec:results_trends}
